---
title: "Falcon Airlines - Capstone Project - Notes 1"
author: "Mohanned Gomaa"
date: "5/10/2020"
output: pdf_document
---

Introduction:
This is the dilemma of a reputed US airline carrier ‘Falcon airlines’. They aim to determine the relative importance of each parameter with regards to their contribution to passenger satisfaction. 

Setting aa working directory 
```{r}
setwd('D:/UT Autin/Capstone/Project Notes I')
getwd()
```
Import load data sets 
```{r}
Flight_data= read.csv("Marketing Project-Flight data.csv",header = T)
Survey_data= read.csv("Marketing Project-Survey data.csv",header = T)
```

Quick Glance over datasets
```{r}
colnames(Flight_data)
colnames(Survey_data)
head(Flight_data)
head(Survey_data)
```

Understand data types we are working with
```{r}
str(Flight_data)
str(Survey_data)
```

Summary overview of the data sets
```{r}
summary(Flight_data)
summary(Survey_data)
class()
```

Another detailed look using the describe function, which shows us the skew of data.
```{r}
library(psych)#Understanding Data distribution and dispersion 
extract1 = describe(Flight_data)
extract2 = describe(Survey_data)
library(openxlsx)
write.xlsx(extract1, file="Flight_data_desc.xlsx", asTable = FALSE)#extract descriptive stats.
write.xlsx(extract2, file="Survey_data_desc.xlsx", asTable = FALSE)#extract descriptive stats.
```

After looking at the results from my descriptive stats, we could not interpret the survey data. So, we are planning to transform categorical ordinal data into their into original form of data. 
```{r}
replce.text = function(variable){
  library(stringr)
  variable = str_replace_all(variable,c("extremely poor"),"0")
  variable = str_replace_all(variable,c("poor"),"1")
  variable = str_replace_all(variable,c("need improvement"),"2")
  variable = str_replace_all(variable,c("acceptable"),"3")
  variable = str_replace_all(variable,c("good"),"4")
  variable = str_replace_all(variable,c("excellent"),"5")
  variable = str_replace_all(variable,c("very inconvinient"),"0")
  variable = str_replace_all(variable,c("Inconvinient"),"1")
  variable = str_replace_all(variable,c("Convinient"),"3")
  variable = str_replace_all(variable,c("very convinient"),"3")
  as.integer(variable)
}


check = Survey_data

View(check)

check$Seat_comfort = replce.text(check$Seat_comfort)
check$Departure.Arrival.time_convenient=replce.text(check$Departure.Arrival.time_convenient)
check$Food_drink = replce.text(check$Food_drink)
check$Gate_location = replce.text(check$Gate_location)
check$Inflightwifi_service = replce.text(check$Inflightwifi_service)
check$Inflight_entertainment = replce.text(check$Inflight_entertainment)
check$Online_support = replce.text(check$Online_support)
check$Ease_of_Onlinebooking = replce.text(check$Ease_of_Onlinebooking)
check$Onboard_service = replce.text(check$Onboard_service)
check$Leg_room_service = replce.text(check$Leg_room_service)
check$Baggage_handling = replce.text(check$Baggage_handling)
check$Checkin_service = replce.text(check$Checkin_service)
check$Cleanliness = replce.text(check$Cleanliness)
check$Online_boarding = replce.text(check$Online_boarding)


View(check)
str(check)

```

A re-run for descriptive stats for survey data after transforming data types.
```{r}
extract3 = describe(check)
write.xlsx(extract3, file="Survey_data_desc2.xlsx", asTable = FALSE)#extract descriptive stats.
```

Common Issue with data:
A- We need to check if we have any missing data and decide on how to treat them. 
```{r}
colSums(is.na(Flight_data))
colSums(is.na(Survey_data))
library(rapportools)
colSums(is.empty(Flight_data))
colSums(is.empty(Survey_data))
library(Amelia)
missmap(Flight_data, main = "Missing values vs observed")
```

B- Check if the data is balanced or not?
```{r}
table(check$Satisfaction)
49761/(90917) #My data is split nearly 50-50 between satisfied or not. 
```

Now I need to visualize my data using ggplot. I attempt try to see the relation between using variables through Uni-variate and Bi-variate analysis...

```{r}

library(ggplot2)

Age_hist = ggplot(master_data)+aes(Age)+geom_histogram(binwidth =3, color ="Orange", alpha = 0.5)+ggtitle(label = 'Passengers Age profile')+ labs(x='Age', y='Passengers')
Age_hist 

flightDist_hist = ggplot(master_data)+aes(master_data$Flight_Distance)+geom_histogram(binwidth =100, color ="Orange", alpha = 0.5)+ggtitle(label = 'Flights Distance profile')+ labs(x='Distance', y='Flights')
flightDist_hist 

DepDelay_hist = ggplot(master_data)+aes(master_data$DepartureDelayin_Mins)+geom_histogram(binwidth =20, color ="Orange", alpha = 0.5)+ggtitle(label = 'Departure Dealy Freq')+ labs(x='Delay in Minutes', y='Freq')

DepDelay_hist

ArrDelay_hist = ggplot(master_data)+aes(master_data$ArrivalDelayin_Mins)+geom_histogram(binwidth =20, color ="Orange", alpha = 0.5)+ggtitle(label = 'Arrival Dealy Freq')+ labs(x='Delay in Minutes', y='Freq')

ArrDelay_hist

SeatComf_hist = ggplot(check)+aes(check$Seat_comfort)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Seat Comfort')+ labs(x='Rating', y='Passengers')
SeatComf_hist

Dep_Arr_time_hist = ggplot(check)+aes(check$Departure.Arrival.time_convenient)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Departure and Arrival Times')+ labs(x='Rating', y='Passengers')
Dep_Arr_time_hist

fB_hist = ggplot(check)+aes(check$Food_drink)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Food & Beverage Ratings')+ labs(x='Rating', y='Passengers')
fB_hist

GL_hist = ggplot(check)+aes(check$Gate_location)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Gate Location Ratings')+ labs(x='Rating', y='Passengers')
GL_hist

wifi_hist = ggplot(check)+aes(check$Inflightwifi_service)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'On-board Wifi Ratings')+ labs(x='Rating', y='Passengers')

wifi_hist

Ent_hist = ggplot(check)+aes(check$Inflight_entertainment)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'On-board Entertainment Ratings')+ labs(x='Rating', y='Passengers')


Ent_hist

OS_hist = ggplot(check)+aes(check$Online_support)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Online Support Ratings')+ labs(x='Rating', y='Passengers')

OS_hist

OB_hist = ggplot(check)+aes(check$Ease_of_Onlinebooking)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Online Booking Ratings')+ labs(x='Rating', y='Passengers')

OB_hist

Ser_hist = ggplot(check)+aes(check$Onboard_service)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Onboard Service Ratings')+ labs(x='Rating', y='Passengers')


Ser_hist

Leg_hist = ggplot(check)+aes(check$Leg_room_service)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Leg Room Ratings')+ labs(x='Rating', y='Passengers')


Leg_hist

bag_hist = ggplot(check)+aes(check$Baggage_handling)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Baggage Handling')+ labs(x='Rating', y='Passengers')


bag_hist


CS_hist = ggplot(check)+aes(check$Checkin_service)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Check-in Service')+ labs(x='Rating', y='Passengers')


CS_hist


Clean_hist = ggplot(check)+aes(check$Cleanliness)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Cleanliness')+ labs(x='Rating', y='Passengers')


Clean_hist

OBB_hist = ggplot(check)+aes(check$Online_boarding)+geom_histogram(binwidth =1, color ="Orange", alpha = 0.5)+ggtitle(label = 'Online Boarding')+ labs(x='Rating', y='Passengers')


OBB_hist

sat_bar = ggplot(check)+aes(check$Satisfaction)+geom_bar( color ="Orange", alpha = 0.5)+ggtitle(label = 'Passengers Overall Satisfaction')+ labs(x='Rating', y='Passengers')


sat_bar

sat_sex_bar = ggplot(master_data)+aes(master_data$Satisfaction)+geom_bar( color ="Orange", alpha = 0.5)+facet_wrap(master_data$Gender~.)+ggtitle(label = 'Passengers Overall Satisfaction by Gender')+ labs(x='Rating', y='Passengers')


sat_sex_bar

sat_Tr_bar = ggplot(master_data)+aes(master_data$Satisfaction)+geom_bar( color ="Orange", alpha = 0.5)+facet_wrap(master_data$TypeTravel~.)+ggtitle(label = 'Passengers Overall Satisfaction by Travel Type')+ labs(x='Rating', y='Passengers')

sat_Tr_bar


sat_class_bar = ggplot(master_data)+aes(master_data$Satisfaction)+geom_bar( color ="Orange", alpha = 0.5)+facet_wrap(master_data$~.)+ggtitle(label = 'Passengers Overall Satisfaction by Class')+ labs(x='Rating', y='Passengers')

sat_class_bar


hist(Car.data$Age[Car.data$Transport == 'Car'], col = 'Grey')

```

```{r}
library(openxlsx)
library(corrplot)
corrplot(cor( master_data_2[,c(-1:-3,-5,-6,-10)]),method = 'square', title = 'Correlation matrix')
Corr = cor(master_data_2[,c(-1:-3,-5,-6,-10)], method ="pearson")
write.xlsx(Corr, file="Corr.xlsx", asTable = FALSE)#extract descriptive stats.
??write.xlsx
```
View(master_data_2)

```

Create a master data set by Join Data sets using customer ID.
```{r}
library(tidyverse)#importing dplyr to use its rename function 

colnames(Flight_data)
colnames(Survey_data)


Survey_data = Survey_data %>% rename(CustomerID = CustomerId) #Renaming cols to prepare for the join
check = check %>% rename(CustomerID = CustomerId) #Renaming cols to prepare for the join

head(Flight_data)

head(Survey_data)

master_data = merge(Flight_data,Survey_data,by.y ='CustomerID') #Consolidating my sets of data into a master data set, using an inner join to improve the analysis and analysis. 

master_data_2 = merge(Flight_data,check,by.y ='CustomerID')
?merge
head(master_data)
```